---
title: "Final Report"
date: last-modified
date-format: "MMMM D, YYYY"
author: 
    - name: Maya Doitch
      affil-id: 1
    - name: Ruben Jimenez
      affil-id: 1
    - name: Alea Seifert
      affil-id: 1
affiliations:
    - id: 1
      name: Department of Statistics, Cal Poly - SLO
abstract: "hqwvdfiqef"
format: 
  pdf:
    number-sections: true
    template-partials: 
      - title.tex
    include-in-header:
      text: |
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
execute: 
  warning: false
  message: false
  echo: false
bibliography: references.bib
---

```{r}
#| label: setup
library(tidyverse)
library(ggtext)
```

```{r}
#| label: data-cleaning
# Data Cleaning


```

```{r, results='hide'}
#| label: data-analysis
 
```

## Introduction {#intro}

A few paragraphs that contain background information, motivation for your\
research, and a statement of your research goals. Be sure to incorporate any supporting\
references into the text. The purpose of the background is to place your work in the greater context of the literature in the area you are investigating. Then you should explicitly identify a hypothesis that you will investigate with your analysis. Don’t assume the reader remembers any of your earlier project reports.

## Data and Methods

### Data Description

-   Briefly describe your data, where your data came from (source), definitions of important variables, and how the data were collected. You should include a variables table and/or diagram representing your study.

-   Indicate any modifications made to the data, recoding, or decisions about missing data

### Statistical Analysis

-   Briefly describe the methods you used (e.g., multilevel regression) in your analysis

-   Do not report results in this section!

## Results {#results}

```{r}
#| label: tbl
#| tbl-cap: ""
#| tbl-pos: "H"


```

@fig

```{r}
#| label: fig
#| fig-cap: "The"
#| fig-align: center
#| fig-pos: "H"


  
```

@fig-cl (X2 = 14.991; df = 1; p-value \< 0.0001).

-   Summarize the results of your exploratory data analysis (e.g., include a matrix plot?). Which Level 1 and Level 2 variables appeared to be most promising before you built your final model

-   Summarize the results of your initial anova exploring the significance of the Level 2 grouping variable.

-   Summarize, including (with a useful picture), your null model and what you learn, including interpreting the ICC value in context.

-   Brief discussion of how you arrived at your final model. Interpret all of the estimated parameters in your final model in context. Include at least one confidence interval or interval of random slopes and interpret in context. (But there shouldn’t really be much “commentary” in this section.) How much of the unexplained Level 1 and/or Level 2\
    variation in the null model is explained by your final model?

-   Integrate the most relevant figures to help interpret the results with your discussion. This includes graphs illustrating fixed effects, random effects, and interactions.

-   These tables and figures should be well-labeled, numbered (e.g., Figure 1), and\
    have a good, descriptive caption.

-   Especially effective graphs compare your data to the model and discuss how the\
    model does and does not capture important features of the data. At least\
    consider showing the “effects plots” in addition to the raw data graphs.

-   Each report should have a minimum of two plots (could be several more, just\
    keep them well-sized and well-integrated into the discussion).

-   Diagnostic analysis: Discussion and inclusion of residual plots (with commentary) for your final model

## Discussion {#discussion}

Describe how the results help answer your research questions and what was most\
interesting from your analysis. In fact, the first paragraph of the Discussion is very\
important – in professional journals, it is often the first and sometimes the only\
paragraph that is read in a paper. After the first sentence highlights primary results, the\
remainder of the first paragraph might compare your results to others in the literature or include interesting secondary results.

-   Discuss possible implications of the results in the context of the research question.

-   Identify any limitations of your study. Discuss the potential impact of such limitations on\
    the conclusions. (e.g., potential confounding variables, generalizability - Don’t give generic statements of possible causation and generalizability, but thoughtfully discuss\
    relevant issues – confounding variables, representativeness of the sample, etc.)

-   Identify strengths and weaknesses of your analysis.

-   Make suggestions for future research. Identify important next steps that a researcher\
    could take to build on your work.

{{< pagebreak >}}

## References {.unnumbered}

```{}
```

```{}
```

```{}
```

```{}
```

```{}
```

```{}
```

## 
