---
title: "Project Part 3"
format: pdf
editor: source
self-contained: true
---

```{r}
#| label: load in data & libraries
set.seed(7000)

library(readr)
library(tidyverse)
library(lme4)
library(performance)
library(ggplot2)
library(ICC)

df_baseball <- read_csv("https://www.dropbox.com/scl/fi/2bcvc8eabdinum2e3r9oj/statcast_pitch_swing_data_20240402_20240630.csv?rlkey=bl9kxe5o9yv017cmjssbgya3s&st=pnrx2dej&dl=1")
```

```{r}
baseball_sample <- df_baseball_clean[sample(nrow(df_baseball_clean), 1000), ] |> arrange(desc(batter))
head(baseball_sample)
```


Include the following modeling steps. This may not find the best model, but will be an opportunity for you to build a multilevel model in a coherent fashion. You should be using your cleaned data set with
quantitative variables grand-mean centered.

1. Include a graph exploring the variability in the response variable across the Level-2 units. Fit an ANOVA using OLS for your response variable and the Level 2 grouping variable (the Level 2 units). Does the variation in the response across the Level 2 units appear to be statistically significant?


```{r}
balanced_sample <- df_baseball_clean |>
  group_by(batter) |>
  filter(n() >= 100) |>
  slice_sample(n=100) |>
  ungroup()
```


```{r}
#| label: graph of variability between top 10 batters (number of hits)

# pull out top 10 hitters by total number of hits
top_batters <- baseball_sample |>
  group_by(batter) |>
  summarize(total_hits = n()) |>
  top_n(10, total_hits) |>
  pull(batter)
filtered_data <- balanced_sam |>
  filter(batter %in% top_batters)

# plot variability in hit distance
ggplot(filtered_data, aes(x = batter, y = hit_distance_sc)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Distance Hit by Batter",
       x = "Batter",
       y = "Distance (ft)") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
model0 <- lm(hit_distance_sc ~ batter, data=balanced_sample)
anova(model0)
```


Based on the results of an ANOVA fitting hit distance by batter, we have strong evidence that batter explains a significant amount of variation in hit distance (F = 2.81, p < .0001).

2. Fit the “random intercepts only” (null) model. Interpret each of the estimated parameters in context. Interpret the intraclass correlation coefficient in context. Does the value of the ICC seem “substantial” to you? Report the likelihood, deviance, and AIC values for later comparison.

```{r}
nullmodel <- lmer(hit_distance_sc ~ 1 + (1 | batter), data=balanced_sample)
summary(nullmodel)

ICC::ICCbare(y=balanced_sample$hit_distance_sc, x = balanced_sample$batter)

logLik(nullmodel)
performance(nullmodel)
```

$$ \tau_0^2: $$ The batter to batter variance in average hit distance is 326.3.

$$ \sigma^2: $$ The variance in average hit distance for each batter is 18029.

$$ \beta_0: $$ The average hit distance across all batters is 164.739.

$$ ICC: \frac{326.3}{326.3 + 18029} = 0.018 \implies$$ The correlation between two hits by the same batter is .018. This is not substantial. The log likelihood of the null model is -181496.3. The deviance is 134.272 feet. The AIC is 363000. 


3. Add 1-3 Level 1 variables. Carry out a likelihood ratio test to compare this model to the model in step 2 (using ML, clearly explain how you find the chi-square value and df). Include details. Also report/compare the AIC values to the intercepts only model. Calculate a “proportion of variation
explained” for this set of variables and interpret the results in context (be clear variation in what). Did the Level 2 variance decrease? What does the tell you? Remove (one at a time) any insignificant variables.

```{r}
#| label: adding level 1 variables

# dist by launch angle, pitch type, launch speed, random batter intercepts
model1 <- lmer(hit_distance_sc ~ launch_angle_gmc + pitch_type + launch_speed_gmc + (1 | batter), data=balanced_sample)
summary(model1)
anova(nullmodel, model1)


# dist by launch angle, launch speed, random batter intercepts
model2 <- lmer(hit_distance_sc ~ launch_angle_gmc + launch_speed_gmc + (1 | batter), data=balanced_sample)
summary(model2)
anova(model2, model1)

# dist by launch angle, random pitch type intercepts, launch speed, random batter intercepts
model3 <- lmer(hit_distance_sc ~ launch_angle_gmc + launch_speed_gmc + (1 | pitch_type) + (1 | batter), data=balanced_sample)

summary(model3)
anova(model3, model1)
```



4. Add 1-3 Level 2 variables. Carry out a likelihood ratio test to compare the models (using ML). Include details. Also report/compare the AIC values. Calculate a “proportion of variation explained” for each level and interpret the results in context. Remove (one at a time) any insignificant
variables.

5. Consider random slopes for one Level 1 variable. (This could involve putting back in one of the variables that was removed earlier...) Include a graph illustrating variability in the estimated random slopes and discuss what you learn in context. Interpret the amount of group-to-group variation in these slopes in context. Once you have a model with at least one set of random slopes, compare this model to the model in step 4, is adding random slopes a significant improvement (REML, be clear how you are determining degrees of freedom)?

6. Add and interpret a cross-level interaction (you may have to use insignificant variables, focus on interpreting the interaction). Are you able to explain much of the slope variation you found in step 5? Is this a significantly better model? Keep in mind: Doing what I tell you to do is ~ B work. Doing more or less will move your grade up or down. Possible Extras: Enhanced graphs; More than 2 levels; Compare model in step 3 to a random effects ANCOVA model (using OLS); Testing additional random slopes; Cross validation (or at least consider possible multiple comparison issues); Including and interpreting confidence intervals